---
title: "p8105_hw2_cl4946.Rmd"
author: "Chen Li"
date: "`r Sys.Date()`"
output: github_document
---
#Problem 1
### To import data and clean names
```{r}
library(tidyverse)

Subway_df = 
  read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")

Subway_df = 
  janitor::clean_names(Subway_df)

```

### Then read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.

```{r}
Subway_df =
  Subway_df |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )

dim(Subway_df)


```
Now this dataset contains subway line and station name, station latitude/ longitude, routes served, entry status, vending availability, entrance type, and ADA. I cleaned the variable names using janitor::clean_names(), selected the variables relevant to this analysis, and converted the entry variable from "YES" and "NO" character values to a logical variable. The resulting dataset has 1868 rows and 19 columns. The data are not fully tidy because route information is spread across route1 to route11 instead of being stored in one variable.



```{r}
###How many distinct stations?
Subway_df |>
  distinct(line, station_name) |>
  nrow()

###How many stations are ADA compliant?

Subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()

###What proportion of station entrances / exits without vending allow entrance?
Subway_df |>
  filter(vending == "NO") |>
  summarize(
    proportion = mean(entry)
  )

### How many distinct stations serve the A train? and how many are ADA compliant?
Subway_long =
  Subway_df |>
  mutate(
    across(route1:route11, as.character)
  ) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name"
  )

Subway_long |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()

Subway_long |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```

# Question 2
### specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel

```{r}
trash_df = 
  readxl::read_excel("./data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N549"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    sports_balls = as.integer(round(sports_balls))
  )


```


### To combine data from from Mr. Trash Wheel, Professor Trash Wheel , and Gwynnda



```{r}
trash_path = "data/202610 Trash Wheel Collection Data.xlsx"

# Mr. Trash Wheel
mr_trash_df =
  readxl::read_excel(
    trash_path,
    sheet = "Mr. Trash Wheel",
    range = readxl::cell_limits(
      c(2, 1),
      c(NA, 14)
    ),
    na = c("", "NA", "dive")
  ) |>
  janitor::clean_names() |>
  tidyr::drop_na(dumpster) |>
  dplyr::mutate(
    dumpster = as.integer(dumpster),
    month = stringr::str_to_title(
      stringr::str_trim(month)
    ),
    year = as.integer(year),
    sports_balls =
      as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel"
  )


# Professor Trash Wheel
professor_trash_df =
  readxl::read_excel(
    trash_path,
    sheet = "Professor Trash Wheel",
    range = readxl::cell_limits(
      c(2, 1),
      c(NA, 13)
    ),
    na = c("", "NA", "dive")
  ) |>
  janitor::clean_names() |>
  tidyr::drop_na(dumpster) |>
  dplyr::mutate(
    dumpster = as.integer(dumpster),
    month = stringr::str_to_title(
      stringr::str_trim(month)
    ),
    year = as.integer(year),
    trash_wheel = "Professor Trash Wheel"
  )


# Gwynnda
gwynnda_df =
  readxl::read_excel(
    trash_path,
    sheet = "Gwynnda the Good Wheel of the W",
    range = readxl::cell_limits(
      c(2, 1),
      c(NA, 12)
    ),
    na = c("", "NA", "dive")
  ) |>
  janitor::clean_names() |>
  tidyr::drop_na(dumpster) |>
  dplyr::mutate(
    dumpster = as.integer(dumpster),
    month = stringr::str_to_title(
      stringr::str_trim(month)
    ),
    year = as.integer(year),
    trash_wheel = "Gwynnda"
  )

trash_df =
  dplyr::bind_rows(
    mr_trash_df,
    professor_trash_df,
    gwynnda_df
  ) |>
  dplyr::select(
    trash_wheel,
    everything()
  )

dim(trash_df)
names(trash_df)
```
###To Calculate the weight: 

```{r}
professor_df =
  trash_df |>
  filter(
    trash_wheel == "Professor Trash Wheel"
  )

professor_weight = 
  professor_df |>
  summarize(
    total_weight = sum(
      weight_tons,
      na.rm = TRUE)
  )

professor_weight =
  professor_weight |>
  pull(total_weight)
```

### And now we want to solve What was the total number of cigarette butts collected by Gwynnda in June of 2022?

```{r}
gwy_2022_df = 
  trash_df |>
  filter(
    trash_wheel == "Gwynnda",
    year == 2022,
    month == "June"
  )

gwy_cig = 
  gwy_2022_df |>
  summarize (
    total_cig = sum(
       cigarette_butts,
    na.rm = TRUE
    )
   
  )

gwy_cig = 
  gwy_cig |>
  pull(total_cig)
```

After cleaning and combining the three datasets, `trash_df` has `r nrow(trash_df)` observations. Some important variables are the Trash Wheel name, dumpster number, month, year, weight of trash, and number of cigarette butts. Professor Trash Wheel collected a total of `r professor_weight` tons of trash. Gwynnda collected `r gwy_cig` cigarette butts in June 2022.

### Question 3
```{r}
mci_baseline_df = 
  read_csv(
    "data/data_mci/MCI_baseline.csv",
    skip =1,
    na = c(".","NA","")
  )|>
  janitor::clean_names()
```
```{r}
names(mci_baseline_df)
```

```{r}
head(mci_baseline_df)
dim(mci_baseline_df)
```
```{r}
mci_baseline_df|>
  mutate(
    sex = case_match(
      sex,
      0 ~ "female",
      1 ~ "male"
    ),
    apoe4 = case_match(
      apoe4,
      0 ~ "non-carrier",
      1 ~ "carrier"
    )
  )
mci_baseline_df |>
  count(sex)
mci_baseline_df |>
  count(apoe4)
```
```{r}
### delete data that don't fit to inclusion criteria

mci_baseline_df = 
  mci_baseline_df |>
  filter(
    is.na(age_at_onset) |
      age_at_onset> current_age
  )

dim(mci_baseline_df)
```
### How many participants were recruited, and of these how many develop MCI? What is the average baseline age? What proportion of women in the study are APOE4 carriers?

```{r}
### How many participants were recruited
n_participants =
  nrow(mci_baseline_df)

n_participants
```

```{r}
### and of these how many develop MCI
developed_mci_df =
  mci_baseline_df |>
  filter(
    !is.na(age_at_onset)
  )
count_developed = 
  nrow(developed_mci_df)

count_developed
```

```{r}
### What is the average baseline age?
average_baseline_age =
  mci_baseline_df |>
  summarize(
    average_age =
      mean(current_age, na.rm = TRUE)
  ) |>
  pull(average_age)

average_baseline_age
```

```{r}
### what is the proportion of women 


women_df =
  mci_baseline_df |>
  filter(
    sex == "female"
  )

women_apoe4_proportion =
  women_df |>
  summarize(
    proportion =
      mean(apoe4 == "carrier")
  ) |>
  pull(proportion)

women_apoe4_proportion
```
There were `r n_participants` participants in the study, and `r count_developed` of them developed MCI during the follow-up period. The average baseline age was `r round(average_baseline_age, 2)` years. Among the women in the study, `r women_apoe4_proportion * 100`% were APOE4 carriers.


### Similarly, import, clean, and tidy the dataset of longitudinally observed biomarker values; comment on the steps on the import process and the features of the dataset.

### Check whether some participants appear in only the baseline or amyloid datasets, and comment on your findings. Combine the demographic and biomarker datasets so that only participants who appear in both datasets are retained, and briefly describe the resulting dataset; export the result as a CSV to your data directory.

```{r}
amyloid_df =
  read_csv(
    "data/data_mci/mci_amyloid.csv",
    skip = 1,
    na = c("NA", "Na", "")
  ) |>
  janitor::clean_names()
```
```{r}
names(amyloid_df)
```

```{r}
dim(amyloid_df)
```

